### 2020-01-29 Manuscript_short.Rmd
Steve's comment addressed, changes in figures, STOPPING POINT VERSION BEFORE CARVING OUT THE NATURAL HISTORY

# Shaped to Kill: The Evolution of Siphonophore Tentilla as Tools for Prey Capture {-}

@@

As predators evolve to feed on different prey taxa, their apparatus for prey capture can adapt into a variety of forms. The study of this evolutionary process is facilitated by a predator clade with structures used exclusively for prey capture and with significant morphological variation in these structures. Siphonophores, a clade of colonial cnidarians, satisfy these criteria particularly well. Their tentilla (tentacle side branches) are the exclusive means of prey capture for the large majority of species and have no other known function. Earlier work has shown that extant siphonophore diets correlate with the different morphologies and sizes of their tentilla and nematocysts. We hypothesize that evolutionary specialization on different prey types has driven the phenotypic evolution of these characters. To test this hypothesis, we: (1) measured multiple morphological characters from siphonophore tentacle specimens from 45 species, (2) mapped these data to a phylogenetic tree, and (3) analyzed the evolutionary associations between morphological characters and prey type data from the literature. Our results show that the siphonophore tentillum morphology has strong evolutionary associations with prey type, and suggest that shifts between prey type are linked to shifts in tentillum and nematocyst size and shape. We found that predatory specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. When there are changes in trophic niche, both trait optima and trait correlation patterns showed significant shifts.  The evolutionary history of tentilla shows that siphonophores are an example of ecological niche diversification via morphological innovation and evolution. The extreme modularity of tentilla may have released siphonophores from the evolutionary constraints of adaptation to ecologically specialized niches. This contributes to understanding how morphological evolution has shaped present-day oceanic food webs.

@@

  Most animal predators use specific structures to capture and subdue prey. Raptors have claws and beaks, snakes have fangs, wasps have stingers, and cnidarians have nematocyst-laden tentacles. The functional morphology of these structures is critical to their ability to successfully capture specific prey [@schmitz2017predator]. Long-term adaptive evolution in response to the defense mechanisms of the prey (*e.g.*, avoidance, escape, protective barriers) leads to modifications that can counter those defenses. The more specialized the diet of a predator is, the more specialized its structures need to be to efficiently overcome the specific challenges posed by the prey. Understanding the relationships between predatory and morphological specializations is necessary to contextualize the phenotypic diversity of predators, quantify the importance of ecological diversification in generating this diversity, and to understand the organismal determinants of food web structure. However, for many clades of predators, there is scarce knowledge on how these specializations evolved with each other. The primary questions we set out to answer are: how do predator specialists and generalists evolve, and how does predatory specialization shape morphological evolution?

  Siphonophores (Cnidaria: Hydrozoa) are a clade of organisms bearing modular structures that are exclusively used for prey capture: the tentilla (Fig. \@ref(figure1)). The tentilla have great morphological variation across species [@mapstone2014global] (Fig. \@ref(figure2)), which makes them an ideal system to study the relationships between functional traits and prey specialization. Like a head of coral, a siphonophore is a colony bearing many feeding polyps (Fig. \@ref(figure1)). Each siphonophore polyp has a single tentacle, which bears a series of side branches known as tentilla. Like other cnidarians, siphonophores capture prey with nematocysts, harpoon-like stinging capsules borne within specialized cells known as cnidocytes. Unlike the prey capture apparatus of most other cnidarians, siphonophore tentacles carry their cnidocytes in extremely complex and organized batteries [@skaer1988formation] which are located in their tentilla. While nematocyst batteries and clusters in other cnidarians are simple static scaffolds for cnidocytes, siphonophore tentilla have their own reaction mechanism, triggered upon encounter with prey. When it fires, a tentillum undergoes an extremely fast conformational change that wraps it around the prey, maximizing the surface area of contact for nematocysts to fire on the prey [@Mackie:1987uy]. In addition, some species have elaborate fluorescent and bioluminescent lures on their tentilla to attract prey with aggressive mimicry [@purcell1980influence;@haddock2005bioluminescent;@haddock2015fluorescent].

@@

  Many siphonophore species inhabit the deep pelagic ocean, which spans from ~200m to the oceanic seafloor. This habitat has fairly homogeneous physical conditions and stable zooplankton abundances and composition [@robison2004deep]. With a relatively predictable prey availability, ecological theory would predict competition to inhibit the coexistence of closely-related species unless evolution towards specialization reduces the breadth of each species’ niche [@simpson1944tempo;@hardin1960competitive;@hutchinson1961paradox]. If this prediction holds true, we would expect the prey capture apparatus morphologies of siphonophores to diversify with the evolution of increasing specialization on a variety of prey types in different siphonophore lineages.

  Specialization has been thought to be an evolutionary ‘dead-end’, meaning that specialized lineages are unlikely to evolve into generalists or to shift the resource for which they are specialized [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization]. However, recent studies have found that interspecific competition can favor the evolution of resource generalism [@stireman2005evolution;@johnson2009competition;@nosil2002transition] and resource switching [@hoberg2008macroevolutionary;@hardy2014specialization]. In addition to studying relationships with morphology, we seek to identify what evolutionary transitions in trophic niche breadth are prevalent in open ocean tactile predators. To do so, we examine three alternative scenarios on siphonophore trophic specialization: (1) predatory specialists evolved from generalist ancestors; (2) predatory specialists evolved from specialist ancestors which specialized on a different resource, switching their primary prey type; and (3) predatory generalists evolved from specialist ancestors. These scenarios are non-exclusive, and each could apply to different transitions along the siphonophore phylogeny.

@@

  Our work builds upon previous pioneering studies that have explored the relationships between tentilla and diet, and show that siphonophores are a robust system for the study of predatory specialization via morphological diversification. [@purcell1984functions] and [@purcell1988correlation] showed clear relationships between diet, tentillum, and nematocyst characters in co-occurring epipelagic siphonophores for a small subset of extant epipelagic siphonophore species.

  In this study, we characterize the morphological diversity of tentilla and their nematocysts across a broad variety of shallow and deep-sea siphonophore species using modern imaging technologies, we expand the phylogenetic tree of siphonophores by combining a broad taxon sampling of ribosomal gene sequences with a transcriptome-based backbone tree, and we explore the evolutionary histories and correlations among diet, tentillum, and nematocyst characters.

@@

*Novel phylogenetic relationships* – We built upon the published siphonophore transcriptome phylogeny [@munro2018improved] to produce a tree with broader taxonomic representation using ribosomal genes. The topology of our tree recapitulates the resolved nodes in [@dunn2005molecular] and [@munro2018improved]. Only 5 nodes (blue dots in Figure \@ref(figure4)) in the unconstrained tree inference were incongruent with the [@munro2018improved] transcriptome tree, and these were constrained during estimation of the constrained 18S+16S tree inference (Fig. \@ref(figure4)). Moreover, with the unprecedented inclusion of *Stephanomia amphytridis* sequences, our tree reveals a novel phylogenetic relationship between the genus *Erenna* and *Stephanomia*.

![\label{figure4} Bayesian time-tree built from 18S + 16S concatenated sequences. Branch lengths estimated using a relaxed molecular clock. Species names in red indicate replicated representation in the morphology data. Species marked with a red asterisk were recorded using high-speed video. Newly accessioned 16S data was used for species marked with a blue circle. Nodes labeled with Bayesian posteriors (BP). Green circles indicate BP = 1. Blue circles indicate nodes constrained to be congruent with (Munro *et al.* 2018). This phylogeny expands the taxon sampling of the (Munro *et al.* 2018) transcriptome tree and is congruent with it. Tips with black squares indicate the species with transcriptomes used in (Munro *et al.* 2018). Tips with grey squares indicate genus-level correspondence to taxa included in (Munro *et al.* 2018). The main clades are labeled: in black for described taxonomic units, and in grey for operational phylogenetic designations.](Figures_final/figure4.png){height=600px}

  We used the clade nomenclature defined in [@dunn2005molecular] and [@munro2018improved], such as Codonophora to indicate the sister group to Cystonectae, Euphysonectae to indicate the sister group to Calycophorae, Clade A and B to indicate the two main lineages within Euphysonectae. In addition, we define two new clades within Codonophora (Fig. \@ref(figure4)): Eucladophora as the clade containing *Agalma elegans* and all taxa that are more closely related to it than to *Apolemia lanosa*, and Tendiculophora as the clade containing *Agalma elegans* and all taxa more closely related to it than to *Bargmannia elongata*. Eucladophora is characterized by bearing spatially differentiated tentilla with proximal heteronemes and a narrower terminal filament region. The etymology derives from the Greek *eu*+*kládos*+*phóros* for “true branch bearers”. Tendiculophora are characterized by bearing rhopalonemes and desmonemes in the terminal filament, having a pair of elastic strands, and developing proximally detachable cnidobands. The etymology of this clade is derived from the Latin *tendicula* for “snare or noose” and the Greek *phóros* for "carriers".

*Evolutionary associations between diet and tentillum morphology* – Reconstructions of feeding guilds do not recover ancestral diet generalism. None of the transitions in diet are consistent with hypothesis 1. Feeding guild specializations have shifted from an alternative ancestral state at least five times, consistent with instances of supporting hypothesis 2 (specialists evolving to feed on a different resource). Copepod specialization and fish specialization evolved twice, and ostracod specialization evolved at least once. We also recover multiple independent origins of generalism from specialist ancestors (Fig. \@ref(figure5)). Large crustacean specialists evolve into generalists twice independently, consistent with instances of hypothesis 3 (generalists evolving from specialists).

To test whether the measured morphological characters evolved in association with shifts in the feeding ecology, we analyzed the evolutionary history of each character on the phylogeny with the feeding guilds reconstructed on it as hypothetical selective regimes. We fitted and compared alternative evolutionary models for each continuous character.  The models compared were the white noise (WN; non-phylogenetic model that assumes all values come from a single normal distribution with no covariance structure among species), the Brownian Motion (BM) model of neutral divergent evolution [@martins1996phylogenies], the Early Burst (EB) model of decreasing rate of evolutionary change [@harmon2010early], and the Ornstein-Uhlenbeck (OU) model of stabilizing selection around a fitted optimum state [@uhlenbeck1930theory;@butler2004phylogenetic]. The model comparison shows that out of 30 characters, 10 show significantly stronger support for the diet-driven multi-optima multi-rate OU model (SM15). These characters include terminal filament nematocyst size and shape, involucrum length, elastic strand width, and heteroneme number. Most of these characters are found exclusively in Tendiculophora, thus this may reflect processes that could be unique to this subtree. Five characters including cnidoband length, cnidoband shape, and haploneme length show maximal support for a diet-driven single-optimum OU model. The remaining 15 characters support BM (or OU with marginal AICc difference with BM).

@@

   We also tested for correlations between shifts in prey selectivity and morphological characters using phylogenetic linear models. We found that fish selectivity is associated with increased number of heteronemes per tentillum, increased roundness of nematocysts (desmonemes and haplonemes), larger heteronemes, reduced heteroneme/cnidoband length ratios, smaller rhopalonemes, lower haploneme SA/V ratios, and increased size of the cnidoband, elastic strand, pedicle and tentacle widths. Decapod-selective diets were associated with increasing cnidoband size and coiledness, haploneme row number, elastic strand width, and heteroneme number. Copepod-selective diets evolved in association with smaller heteroneme and total nematocyst volumes, smaller cnidobands, rounder rhopalonemes, elongated heteronemes, narrower haplonemes with higher SA/V ratios, and smaller heteronemes, tentacles, pedicles, and elastic strands. Selectivity for ostracods was associated with reductions in size and number of heteroneme nematocysts, reductions in cnidoband size, number of haploneme rows, heteroneme numbers, and cnidoband coiledness. Heteroneme length and elongation also correlated negatively with chaetognath selectivity (SM21). These results indicate that not only diet but also differential feeding selectivity has evolved in correlation with changes in the prey capture apparatus of siphonophores.

@@

*Evolution of siphonophore trophic niche* – Siphonophores are an abundant group of zooplankton in oceanic ecosystems [@longhurst1985structure;@o2007copepod]. While little is known about siphonophore trophic ecology, what is known indicates that they occupy a central position in midwater food webs [@choy2017deep], serving as important trophic intermediaries between smaller zooplankton and higher trophic level predators. Siphonophore species have been observed to feed on a variety of prey with very different sizes, traits, and behaviors. Because there is a total absence of siphonophores in the fossil record, how they became established as the ubiquitous and trophically diverse predators in today’s oceans remains an open question.

According to our reconstructions, the evolutionary history of siphonophore diets indicates that being a specialist was an ancestral aspect of their trophic niche, while trophic generalism is likely a derived condition. Several studies [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization] have suggested that resource specialization is an irreversible dead-end due to the constraints posed by extreme phenotypic specialization. Our results show that this is not the case for siphonophores, where the prey type on which they specialize has shifted at least 5 times, and generalism has evolved independently from specialists at least twice. Among the evolutionary scenarios considered, we find support for both scenario 2 (specialist resource switching) and 3 (specialist to generalist), but no support for scenario 1 (generalist to specialist). This finding is particularly given in that it is the opposite of known technical biases in ancestral state reconstruction. [@nosil2005testing] found that such methods tend to infer higher transition rates toward the more frequent state. In this case, that would lead to a bias for an increased rate from generalists (the rarer state across the tips) to specialists (the more common state across the tips). We observe the opposite, indicating strong evidence that these generalists are indeed a derived state.

@@

*Heterochrony and convergence in the evolution of tentilla with diet* - Copepod-specialized diets have evolved convergently in *Cordagalma* and some calycophorans. These evolutionary transitions happened together with transitions to smaller tentilla with fewer cnidoband nematocysts. Tentilla are expensive single-use structures [@Mackie:1987uy], therefore we would expect that specialization in small prey would beget reductions in the size of the prey capture apparatus to the minimum required for the ecological performance. *Cordagalma*’s tentilla strongly resemble the larval tentilla (only found in the first-budded feeding body of the colony) of their sister genus *Forskalia*. This indicates that the evolution of *Cordagalma* tentilla could be a case of paedomorphic heterochrony associated with predatory specialization on smaller prey.

@@

*Limitations* - While our results unambiguously show that tentillum morphology evolved with diet, the conclusions we can draw from these analyses are limited in several ways. The biggest challenge at present is the sparse dietary data available in the literature. Moreover, our analyses are not sufficient to test hypotheses of adaptation. When interpreting these results, it is also important to remember that diet is also dependent on environmental prey availability. Selectivity differences across siphonophore species could be also driven by other phenotypes not accounted for in this study. Further observations on behavior, digestion biochemistry, and toxin composition are necessary to assess their relative importance in determining diet. Our results strongly support the presence of deviations from the generalist-to-specialist evolution model. However, additional dietary characterization could reveal transitions from generalists to specialists for two reasons. First, some of the taxa in our dataset have a very limited number of feeding observations, which could lead to apparent specialization. Second, some of the taxa not included in our dataset could be undiscovered generalists.

@@

  The implications of these results to the evolution of nematocyst function are that an innovation in the discharge mechanism of haplonemes may have occurred during the main shift to elongation. Elongate nematocysts can be tightly packed into cnidobands. We hypothesize this may be a Tendiculophora lineage-specific adaptation to packing more nematocysts into a limited tentillum space, as suggested by [@skaer1988formation]. [@thomason1988allometry] hypothesized that smaller, more spherical nematocysts, with a lower surface area to volume ratio, are more efficient in osmotic-driven discharge and thus have more power for skin penetration. The elongated haplonemes of crustacean-eating Tendiculophora have never been observed penetrating their crustacean prey ([@purcell1984functions] and our unpublished observations), and are hypothesized to entangle the prey through adhesion of the abundant spines to the exoskeletal surfaces and appendages. Entangling requires less acceleration and power during discharge than penetration, as it does not rely on point pressure. In fish-eating cystonects and *Erenna* species, the haplonemes are much less elongated and very effective at penetration, in congruence with the osmotic discharge hypothesis. Tendiculophora, comprised of the clades Euphysonectae and Calycophorae, includes the majority of siphonophore species. Within these clades are the most abundant siphonophore species, and a greater morphological and ecological diversity is found. We hypothesize that this packing-efficient haploneme morphology may have been a key innovation leading to the diversification of this clade. However, other characters that shifted concurrently in the stem of this clade may have been responsible for their extant diversity.

@@

  Our evolutionary rate covariance results (SM36-38) indicate that tentilla are not only phenotypically integrated but also show patterns of evolutionary modularity, where different sets of characters appear to evolve in stronger correlations among each other than with other characters [@wagner1996homologues]. This may be indicative of the underlying genetic and developmental dependencies among closely homologous nematocyst types (such as desmonemes and rhopalonemes) and structures. In addition, these evolutionary modules point to hypothetical functional modules. For example, the coiling degree of the cnidoband and the extent of the involucrum have correlated rates of evolution, while high-speed videos show that the involucrum helps direct the whiplash of the uncoiling cnidoband forward (towards the prey).

  While selection acting on character states is a widely studied phenomenon, recent studies have shown that selection can also act upon the patterns of character correlations and phenotypic dependencies [@young2005serial;@goswami2006morphological;@revell2009phylogenetic;@monteiro2010adaptive;@hallgrimsson2012generation;@claverie2013modularity;@caetano2018estimating]. This evolution of character relationships can allow lineages to explore new regions of the morphospace and facilitate the appearance of ecological novelties. Our results show that the patterns of phenotypic integration in siphonophore tentilla vary among clades, and appear to display different relationships across shifting feeding specializations.  Similarly to what has been found in the feeding morphologies of fish [@collar2005comparative;@revell2009phylogenetic], siphonophore tentilla may have accommodated new diets by altering the correlations between characters. For example, changes in the size and shape relationships between nematocyst types gave rise to the nematocyst complements specialized in ensnaring prey with different combinations of defensive traits.

>>>>

# Shaped to kill: The evolution of siphonophore tentilla for specialized prey capture in the open ocean {-}

@@

As predators evolve to feed on different prey taxa, their apparatus for prey capture can adapt into a variety of forms. The study of this evolutionary process is facilitated by a predator clade with structures used exclusively for prey capture and with significant morphological variation in these structures. Siphonophores, a clade of colonial cnidarians, satisfy these criteria particularly well. Their tentilla (tentacle side branches) are the exclusive means of prey capture for the large majority of species and have no other known function. Earlier work has shown that extant siphonophore diets correlate with the different morphologies and sizes of their tentilla and nematocysts. We hypothesize that evolutionary specialization on different prey types has driven the phenotypic evolution of these characters. To test this hypothesis, we: (1) measured multiple morphological characters from siphonophore tentacle specimens from 45 species, (2) mapped these data to a phylogenetic tree, and (3) analyzed the evolutionary associations between morphological characters and prey type data from the literature. Our results show that siphonophore tentillum morphology has strong evolutionary associations with prey type, and suggest that shifts between prey type are linked to shifts in tentillum and nematocyst size and shape. We found that predatory specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. When there are changes in trophic niche, both trait optima and trait correlation patterns showed significant shifts.  The evolutionary history of tentilla shows that siphonophores are an example of ecological niche diversification via morphological innovation and evolution. The extreme modularity of tentilla may have released siphonophores from the evolutionary constraints of adaptation to ecologically specialized niches. This contributes to understanding how morphological evolution has shaped present-day oceanic food webs.

@@

  Most animal predators use specific structures to capture and subdue prey. Raptors have claws and beaks, snakes have fangs, wasps have stingers, and cnidarians have nematocyst-laden tentacles. The functional morphology of these structures is critical to their ability to successfully capture prey [@schmitz2017predator]. Long-term adaptive evolution in response to the defense mechanisms of the prey (*e.g.*, avoidance, escape, protective barriers) leads to modifications that can counter those defenses. The more specialized the diet of a predator is, the more specialized its structures need to be to efficiently overcome the challenges posed by the prey. Understanding the relationships between predatory and morphological specializations is necessary to contextualize the phenotypic diversity of predators, quantify the importance of ecological diversification in generating this diversity, and to understand the organismal determinants of food web structure. However, for many clades of predators, there is scarce knowledge on how these specializations evolved with each other. The primary questions we set out to answer are: how do predator specialists and generalists evolve, and how does predatory specialization shape morphological evolution?

  Siphonophores (Cnidaria: Hydrozoa) are a clade of organisms bearing modular structures that are exclusively used for prey capture: the tentilla (Fig. \@ref(figure1)). The tentilla have great morphological variation across species [@mapstone2014global] (Fig. \@ref(figure2)), which makes them an ideal system to study the relationships between functional traits and prey specialization. Like a head of coral, a siphonophore is a colony bearing many feeding polyps (Fig. \@ref(figure1)). Each feeding polyp has a single tentacle, which bears a series of side branches known as tentilla. Like other cnidarians, siphonophores capture prey with nematocysts, harpoon-like stinging capsules borne within specialized cells known as cnidocytes. Unlike the prey capture apparatus of most other cnidarians, siphonophore tentacles carry their cnidocytes in extremely complex and organized batteries [@skaer1988formation] which are located in their tentilla. While nematocyst batteries and clusters in other cnidarians are simple static scaffolds for cnidocytes, siphonophore tentilla have their own reaction mechanism, triggered upon encounter with prey. When it fires, a tentillum undergoes an extremely fast conformational change that wraps it around the prey, maximizing the surface area of contact for nematocysts to fire on the prey [@Mackie:1987uy]. In addition, some species have elaborate fluorescent and bioluminescent lures on their tentilla to attract prey with aggressive mimicry [@purcell1980influence;@haddock2005bioluminescent;@haddock2015fluorescent].

@@

  Many siphonophore species inhabit the deep pelagic ocean, which spans from ~200m to the abyssal seafloor (~4000m). This habitat has fairly homogeneous physical conditions and stable zooplankton abundances and composition [@robison2004deep]. With a relatively predictable prey availability, ecological theory predicts that interspecific competition would inhibit the coexistence of closely-related species unless evolution towards specialization reduces the breadth of each species’ niche [@simpson1944tempo;@hardin1960competitive;@hutchinson1961paradox]. If this prediction holds true, we would expect the prey capture apparatus morphologies of siphonophores to diversify with the evolution of increasing specialization on a variety of prey types in different siphonophore lineages.

  Specialization has been thought to be an evolutionary ‘dead-end’, meaning that specialized lineages are unlikely to evolve into generalists or to shift the resource for which they are specialized [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization]. However, recent studies have found that interspecific competition can favor the evolution of resource generalism [@stireman2005evolution;@johnson2009competition;@nosil2002transition] and resource switching [@hoberg2008macroevolutionary;@hardy2014specialization]. In addition to studying relationships with morphology, we seek to identify what evolutionary transitions in trophic niche breadth are prevalent in open-ocean tactile predators. To do so, we examine three alternative scenarios of siphonophore trophic specialization: (1) predatory specialists evolved from generalist ancestors; (2) predatory specialists evolved from specialist ancestors which targeted different resources, switching their primary prey type; and (3) predatory generalists evolved from specialist ancestors. These scenarios are non-exclusive, and each could apply to different transitions along the siphonophore phylogeny.

@@

  Our work builds upon previous pioneering studies that have explored the relationships between tentilla and diet, and showed that siphonophores are a robust system for the study of predatory specialization via morphological diversification. [@purcell1984functions] and [@purcell1988correlation] showed clear relationships between diet, tentillum, and nematocyst characters in co-occurring epipelagic siphonophores for a small subset of extant epipelagic siphonophore species.

  In this study, we characterize the morphological diversity of tentilla and their nematocysts across a broad variety of shallow and deep-sea siphonophore species using modern imaging technologies, we expand the phylogenetic tree of siphonophores by combining ribosomal gene sequences from a broad range of taxa with a transcriptome-based backbone tree, and we explore the evolutionary histories and correlations between diet, tentillum, and nematocyst characters.

@@

*Novel phylogenetic relationships* – We built upon the published siphonophore transcriptome phylogeny [@munro2018improved] to produce a tree with broader taxonomic representation using ribosomal genes. The topology of our tree recapitulates the resolved nodes in [@dunn2005molecular] and [@munro2018improved]. Only 5 nodes (blue dots in Figure \@ref(figure4)) in the unconstrained tree inference were incongruent with the [@munro2018improved] transcriptome tree, and these were constrained to the [@munro2018improved] topology during estimation of the constrained 18S+16S tree inference (Fig. \@ref(figure4)). Moreover, with the inclusion of *Stephanomia amphytridis* sequences, our tree reveals a novel phylogenetic relationship between the genus *Erenna* and *Stephanomia*.

![\label{figure4} Bayesian time-tree built from 18S + 16S concatenated sequences. Branch lengths estimated using a relaxed molecular clock. Species names in red indicate replicated representation in the morphology data. Species marked with a red asterisk were recorded using high-speed video. Newly accessioned 16S data was used for species with names in bold. Nodes labeled with Bayesian posteriors (BP). Green circles indicate BP = 1. Blue circles indicate nodes constrained to be congruent with Munro *et al.* (2018). Tips with black squares indicate the species with transcriptomes used in Munro *et al.* (2018). Tips with grey squares indicate genus-level correspondence to taxa included in Munro *et al.* (2018). The main clades are labeled: in black for described taxonomic units, and in grey for operational phylogenetic designations.](Figures_final/figure4.png){height=600px}

  We used the clade nomenclature defined in [@dunn2005molecular] and [@munro2018improved], including Codonophora to indicate the sister group to Cystonectae, Euphysonectae to indicate the sister group to Calycophorae, Clade A and B to indicate the two main lineages within Euphysonectae. In addition, we define two new clades within Codonophora (Fig. \@ref(figure4)): Eucladophora as the clade containing *Agalma elegans* and all taxa that are more closely related to it than to *Apolemia lanosa*, and Tendiculophora as the clade containing *Agalma elegans* and all taxa more closely related to it than to *Bargmannia elongata*. Eucladophora is characterized by bearing spatially differentiated tentilla with proximal heteronemes and a narrower terminal filament region. The etymology derives from the Greek *eu*+*kládos*+*phóros* for “true branch bearers”. Tendiculophora are characterized by bearing rhopalonemes and desmonemes in the terminal filament, having a pair of elastic strands, and developing proximally detachable cnidobands. The etymology of this clade is derived from the Latin *tendicula* for “snare or noose” and the Greek *phóros* for "carriers".

*Evolutionary associations between diet and tentillum morphology* – Reconstructions of feeding guilds do not recover ancestral diet generalism. None of the transitions in diet are consistent with scenario 1 (specialists evolving from generalists). Feeding guild specializations have shifted from an alternative ancestral state at least five times, consistent with instances supporting scenario 2 (specialists evolving to feed on a different resource). Copepod specialization and fish specialization evolved twice, and ostracod specialization evolved at least once. We also recover multiple independent origins of generalism from specialist ancestors (Fig. \@ref(figure5)). Large crustacean specialists evolve into generalists twice independently, consistent with instances of scenario 3 (generalists evolving from specialists). This finding is particularly compelling given in that it is the opposite of known biases in ancestral state reconstruction. [@nosil2005testing] found that such methods tend to infer higher transition rates toward the more frequent state. In this case, that would lead to a bias for an increased rate of transition from generalists (the rarer state across the tips) to specialists (the more common state across the tips). We observe the opposite, indicating strong evidence that these generalists are indeed a derived state.

To test whether measured morphological characters evolved in association with shifts in feeding ecology, we analyzed the evolutionary history of each character on the phylogeny, with the feeding guilds reconstructed on it as hypothetical selective regimes. We fitted and compared alternative evolutionary models for each continuous character. The models compared were the white noise (WN; non-phylogenetic model that assumes all values come from a single normal distribution with no covariance structure among species), the Brownian Motion (BM) model of neutral divergent evolution [@martins1996phylogenies], the Early Burst (EB) model of decreasing rate of evolutionary change [@harmon2010early], and the Ornstein-Uhlenbeck (OU) model of stabilizing selection around a fitted optimum state [@uhlenbeck1930theory;@butler2004phylogenetic]. The model comparison shows that out of 30 characters, 10 show significantly stronger support for the diet-driven multi-optima multi-rate OU model (SM15). These characters include terminal filament nematocyst size and shape, involucrum length, elastic strand width, and heteroneme number. Most of these characters are found exclusively in Tendiculophora, thus this may reflect processes that could be unique to this subtree. Five characters including cnidoband length, cnidoband shape, and haploneme length show maximal support for a diet-driven single-optimum OU model. The remaining 15 characters support BM (or OU with marginal AICc difference with BM).

@@

   We also tested for correlations between shifts in prey selectivity and morphological characters using phylogenetic linear models. We found that fish selectivity is associated with increased number of heteronemes per tentillum, increased roundness of nematocysts (desmonemes and haplonemes), larger heteronemes, reduced heteroneme/cnidoband length ratios, smaller rhopalonemes, lower haploneme SA/V ratios, and larger the cnidoband, elastic strand, pedicle and tentacle widths. Decapod-selective diets were associated with increasing cnidoband size and coiledness, haploneme row number, elastic strand width, and heteroneme number. Copepod-selective diets evolved in association with smaller heteroneme and total nematocyst volumes, smaller cnidobands, rounder rhopalonemes, elongated heteronemes, narrower haplonemes with higher SA/V ratios, and smaller heteronemes, tentacles, pedicles, and elastic strands. Selectivity for ostracods was associated with reductions in size and number of heteroneme nematocysts, reductions in cnidoband size, number of haploneme rows, heteroneme numbers, and cnidoband coiledness. Heteroneme length and elongation also correlated negatively with chaetognath selectivity (SM21). These results indicate that not only diet but also differential feeding selectivity has evolved in correlation with changes in the prey capture apparatus of siphonophores.

@@

*Evolution of siphonophore trophic niche* – Siphonophores are an abundant group of zooplankton in oceanic ecosystems [@longhurst1985structure;@o2007copepod]. While little is known about siphonophore trophic ecology, what is known indicates that they occupy a central position in midwater food webs [@choy2017deep], serving as important trophic intermediaries between smaller zooplankton and higher trophic level predators. Siphonophore species have been observed to feed on a variety of prey with very different sizes, traits, and behaviors. Because there is a total absence of siphonophores in the fossil record, how they became established as such ubiquitous and trophically diverse predators in today’s oceans remains an open question.

Several studies [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization] have suggested that resource specialization is an irreversible dead-end due to the constraints posed by extreme phenotypic specialization. Our results show that this is not the case for siphonophores, where the prey type on which they specialize has shifted at least 5 times. We find no support for any transitions from generalist to specialist (scenario 1, as described in the Introduction). We do find support for at least 3 instances of specialists switching from one prey type to another prey type, (scenario 2) and two switches from specialist to generalist (scenario 3).

@@

*Heterochrony and convergence in the evolution of tentilla with diet* - In addition to identifying shifts in prey type, our work reveals the specific morphological changes in the prey capture apparatus associated with these changes. Copepod-specialized diets have evolved convergently in *Cordagalma* and some calycophorans. These evolutionary transitions happened together with transitions to smaller tentilla with fewer cnidoband nematocysts. Tentilla are expensive single-use structures [@Mackie:1987uy], therefore we would expect that specialization in small prey would beget reductions in the size of the prey capture apparatus to the minimum required for the ecological performance. *Cordagalma*’s tentilla strongly resemble the larval tentilla (only found in the first-budded feeding body of the colony) of their sister genus *Forskalia*. This indicates that the evolution of *Cordagalma* tentilla could be a case of paedomorphic heterochrony associated with predatory specialization on smaller prey.

@@

@@

  The implications of these results for the evolution of nematocyst function are that an innovation in the discharge mechanism of haplonemes may have occurred during the main shift to elongation. Elongate nematocysts can be tightly packed into cnidobands. We hypothesize this may be a Tendiculophora lineage-specific adaptation to packing more nematocysts into a limited tentillum space, as suggested by [@skaer1988formation]. [@thomason1988allometry] hypothesized that smaller, more spherical nematocysts, with a lower surface area to volume ratio, are more efficient in osmotic-driven discharge and thus have more power for skin penetration. The elongated haplonemes of crustacean-eating Tendiculophora have never been observed penetrating their crustacean prey ([@purcell1984functions] and our unpublished observations), and are hypothesized to entangle the prey through adhesion of the abundant spines to the exoskeletal surfaces and appendages. Entangling requires less acceleration and power during discharge than penetration, as it does not rely on point pressure. In fish-eating cystonects and *Erenna* species, the haplonemes are much less elongated and very effective at penetration, in congruence with the osmotic discharge hypothesis. Tendiculophora, comprised of the clades Euphysonectae and Calycophorae, includes the majority of siphonophore species. Within these clades are the most abundant siphonophore species, and a greater morphological and ecological diversity is found. We hypothesize that this packing-efficient haploneme morphology may have been a key innovation leading to the diversification of this clade. However, other characters that shifted concurrently in the stem of this clade may have been responsible for their extant diversity.

@@

  Our evolutionary rate covariance results (SM36-38) indicate that tentilla are not only phenotypically integrated but also show patterns of evolutionary modularity, where different sets of characters appear to evolve in stronger correlations among each other than with other characters [@wagner1996homologues]. This may be indicative of the underlying genetic and developmental dependencies among closely homologous nematocyst types (such as desmonemes and rhopalonemes) and structures. In addition, these evolutionary modules point to hypothetical functional modules. For example, the coiling degree of the cnidoband and the extent of the involucrum have correlated rates of evolution, while high-speed videos show that the involucrum helps direct the whiplash of the uncoiling cnidoband distally (towards the prey).

  While selection acting on character states is a widely studied phenomenon, recent studies have shown that selection can also act upon the patterns of character correlations and phenotypic dependencies [@young2005serial;@goswami2006morphological;@revell2009phylogenetic;@monteiro2010adaptive;@hallgrimsson2012generation;@claverie2013modularity;@caetano2018estimating]. This evolution of character relationships can allow lineages to explore new regions of the morphospace and facilitate the appearance of ecological novelties. Our results show that the patterns of phenotypic integration in siphonophore tentilla vary among clades, and appear to display different relationships across shifting feeding specializations.  Similar to what has been found in the feeding morphologies of fish [@collar2005comparative;@revell2009phylogenetic], siphonophore tentilla may have accommodated new diets by altering the correlations between characters. For example, changes in the size and shape relationships between nematocyst types gave rise to the nematocyst complements specialized in ensnaring prey with different combinations of defensive traits.

*Limitations* - While our results unambiguously show that tentillum morphology evolved with diet, and strongly support deviations from the generalist-to-specialist evolution scenario, the conclusions we can draw from these analyses are limited in several ways. The biggest challenge at present is the sparse dietary data available in the literature. Additional dietary data could reveal transitions from generalists to specialists we were unable to detect for two reasons. First, some of the taxa in our dataset have a very limited number of feeding observations, which could lead to apparent specialization. Second, some of the taxa not included in our dataset could be undiscovered generalists. When interpreting these results, it is also important to remember that diet is also dependent on environmental prey availability. In addition, selectivity differences across siphonophore species could be also driven by other phenotypes not accounted for in this study. Finally, further observations on behavior, digestion biochemistry, and toxin composition are necessary to assess their relative importance in determining diet.
